---
title: "Miles per Gallon"
author: "Alexander Alexandrov"
output: pdf_document
mode: selfcontained
---

# Overview

Work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

* "Is an automatic or manual transmission better for MPG?"
* "Quantify the MPG difference between automatic and manual transmissions"

Its evident from my experience that *transimission type* is not the only variable that has impact on *mpg*. So several models should be analized and compared to each over. Following parameters should be considered:

* Transmission type
* Weight
* Gross horsepower
* Number of cylinders
* Number of forward gears

Some of them are correlated. To exclude injurious effect residuals's analisys should be made.

\newpage

# Motor Trend Cars Data Overview

Load Motor Trend cars data. And adjust some variables.

```{r}
library(data.table)
data(mtcars)
data <- data.table(mtcars)
invisible({
    data[, vs := factor(vs)]
    data[, am := factor(am)]
    data[, cyl := factor(cyl)]
    data[, gear := factor(gear)]
});
str(data)
```

Data is clean enough.

|Variable|Type|Domain|Description|
|---|---|---|---|
|**mpg**|numeric|positive real number|Miles/(US) gallon|
|**cyl**|factor|4, 6, 8|Number of cylinders|
|**disp**|numeric|positive real number|Displacement (cu.in.)|
|**hp**|numeric|positive real number|Gross horsepower|
|**drat**|numeric|positive real number|Rear axle ratio|
|**wt**|numeric|positive real number|Weight (lb/1000)|
|**qsec**|numeric|positive real number|1/4 mile time|
|**vs**|factor|0, 1|V/S|
|**am**|factor|0, 1|Transmission (0 = automatic, 1 = manual)|
|**gear**|factor|3, 4, 5|Number of forward gears|
|**carb**|numeric|positive integer number|Number of carburetors|

\newpage

# Further exploratory data analyses

Some usefull plots.

```{r}
library(ggplot2)

```

According to quantiles (boxplots) orange juice (red) is slightly better then ascorbic acid.

\newpage
